\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-05-20 paper.qmd
Added Lwd MSE plot

@@

<!--add distribution of distance along with the default benchmark-->
<!--show the trimesh with colored long edges-->

>>>>

<!--To generate errors for different total number of bins-->
```{r}
## To initialize effective bins along x
effective_bin1_scurve <- 10

## To initialize benchmark values to remove low density hexagons
benchmark_rm_hex_vec <- append(seq(0, 0.99, by=0.1),0.99)

error_rm_scurve <- data.frame(matrix(nrow = 0, ncol = 0))

effective_bin2_scurve <- calc_bins_y(bin1 = effective_bin1_scurve, r2 = r2)$bin2

scurve_model <- fit_highd_model(
  training_data = training_data_scurve,
  emb_df = umap_scurve_scaled,
  bin1 = effective_bin1_scurve,
  r2 = r2,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "x"
)

for (benchmark_rm_lwd in benchmark_rm_hex_vec) {

  df_bin_centroids_scurve_high_dens <- df_bin_centroids_scurve |>
    filter(std_counts > benchmark_rm_lwd)

  df_bin_scurve_high_dens <- df_bin_scurve |>
    filter(hb_id %in% df_bin_centroids_scurve_high_dens$hexID)

  ## Compute error
  error_df <- glance(
    df_bin_centroids = df_bin_centroids_scurve_high_dens,
    df_bin = df_bin_scurve_high_dens,
    training_data = training_data_scurve,
    newdata = NULL,
    type_NLDR = "UMAP",
    col_start = "x") |>
    mutate(benchmark_rm_lwd = benchmark_rm_lwd,
           bin1 = effective_bin1_scurve,
           bin2 = effective_bin2_scurve,
           b = bin1 * bin2,
           b_non_empty = NROW(df_bin_centroids_scurve_high_dens))

  error_rm_scurve <- bind_rows(error_rm_scurve, error_df)

benchmark_label_df <- tibble(x = 14, y = 0.6, 
                          label = "benchmark is 0.60.")

mse_scurve_lwd <- ggplot(error_rm_scurve, 
                     aes(x = b_non_empty, 
                         y = log(MSE))) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = 15, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  geom_text(data = benchmark_label_df, 
            aes(x=x, 
                label= label, 
                y=y), 
            colour="red", 
            angle=90) +
  ylab("log(MSE)") +
  xlab("number of non-empty bins") +
  theme(axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 5),
        axis.title.y = element_text(size = 5, angle = 90))
```

```{r}
#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent bechmark choices. What is the effective bechmark value to remove low density hexagons? The residual plot have a steep slope at the beginning, indicating that a smaller bechmark value causes a larger amount of MSE. Then, the slope gradually declines or level off, indicating that a higher bechmark value generates a smaller MSE. Using the elbow method, it was observed that when the bechmark value is set to $0.6$, the lowest MSE occurred."
#| label: fig-mse-scurve-lwd
#| fig-pos: H
#| out-width: 50%

mse_scurve_lwd
```

@@

<!--add distribution of distance along with the default benchmark-->
<!--show the trimesh with colored long edges with different benchmark values-->

### 2024-05-20 paper.tex
Added Lwd MSE plot

>>>>

\includegraphics[width=0.5\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mse-scurve-lwd-1.pdf}

\caption{\label{fig-mse-scurve-lwd}MSE from UMAP applied to S-curve
dataset with diffferent bechmark choices. What is the effective bechmark
value to remove low density hexagons? The residual plot have a steep
slope at the beginning, indicating that a smaller bechmark value causes
a larger amount of MSE. Then, the slope gradually declines or level off,
indicating that a higher bechmark value generates a smaller MSE. Using
the elbow method, it was observed that when the bechmark value is set to
\(0.6\), the lowest MSE occurred.}
